---
title: "Aides Drug"
author: "Tracy Alset"
date: "07/12/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
female<- filter(triUmp, gender=="F")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

We have to take out the females using the filter function and then run graphical, numerical, and inferential tests to find out if the drug is effective or not.


```{r eval=FALSE}
female<-female%>%
mutate(across(where(is.character), as.factor))

```



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")

```

It looks like out of the females who got infected with aids:
the ones who took the drug were more than the ones who took the placebo.
The ones who weren't infected:
more of them took the placebo than the ones who took the drug.
this kinda indicates that the drug isn't effective really.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)
```

Out of the females that were infected with aids:
the higher percentage of infection was for the ones that took the drug , and less for the ones who took the placebo. this shows that the drug wasn't effective.
the second table doesn't show data that makes sense, because it doesn't any data that shows the effectiveness of the drug.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

We reject the null hypothesis because of the low P-value , and the null hypothesis said that there is no difference and that is rejected so there is a difference between the females who took the drug and the ones who didn't take it (meaning the drug isn't effective).


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->






